Kaggle-in-class Data Challenges Can Boost Student Learning

### 2021-01-06 paper-kaggle.Rmd
ecdf plots plus some typos

<!-- Blinded verssion -->
The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of AB (Computational statistics and data mining ,MAST90083, denoted as CSDM), and one at CD University (Statistical thinking, ETC2420/5242, denoted as ST).

@@

\end{tabular}
\caption{Computational statistics and data mining: Summary statistics of the exam score (out of 100) and the second assignment (out of 10) for the two competition groups.}
\end{center}
\label{tab:Melb_Group_Compare}
\end{table}

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{l l | r | r | r r r r r}\hline
Score & Competition group & N & Mean (SD)   & Min  &  Q1 & Median  &  Q3  & Max \\\hline
Exam   & ST-PG      & 34  & 61.8 (12.2)  &  35.0 & 57.2  &   61.0  & 72.2  & 81.5 \\
       & ST-UG         & 145 & 63.0 (15.4)  &   0.0 & 56.0  &   64.0  & 73.0   & 90.0\\\hline
\end{tabular}
\caption{Statistical thinking: Summary statistics of the exam score (out of 100) for the two groups.}
\end{center}
\label{tab:Melb_Group_Compare}
\end{table}

@@

```{r Monash_quiz, fig.cap = "\\label{fig:Monash_quiz}Beeswarm plots of undergraduate (ST-UG) and postgraduate (ST-PG) students quizzes during the semester. The scores are very similar, although undergraduates have a longer lower tail.", fig.height=6, fig.width=6, eval=FALSE}
library(tidyverse)
library(gridExtra)
library(ggbeeswarm)
library(forcats)
quiz <- read_csv(file="data/quiz_scores_by_unit_2420_5242.csv")

@@

ggplot(quiz, aes(x=unit, y=score, colour=unit)) +
  geom_beeswarm(alpha=0.5) +
  facet_wrap(~quiz) + 
  xlab("") + ylab("Score (%)") +
  theme(legend.position="none")

@@

ggplot(quiz, aes(x=score^4, colour=unit, fill=unit)) +
  geom_density(alpha=0.5) +
  facet_grid(unit~quiz, scales="free") + 
  xlab("") + ylab("Score (%)") +
  theme(legend.position="none")

@@

```{r ETC5242, fig.cap = "\\label{fig:ETC5242} Performance for regression question relative to total exam score for students who did and didn't do the regression data competition in Statistical thinking.", fig.width=4, fig.height=3}
scores_5242 <- read_csv("data/deidentified_scores_5242.csv")
scores_2420 <- read_csv("data/deidentified_scores_2420.csv")

@@

Because the experiment was conducted in the classroom setting as part of the normal teaching of the courses, which imposed limitations on the design. However, performance comparison was enabled in CSDM by a randomized assignment of students to two topic groups, and in ST by using a comparison group. In CDSM, the group sizes were relatively small, approximately 30 students per group. For ST the control group was the undergraduate students that took class. Using undergraduate students as a control group for graduate students may be surprising.  However, the experience of teaching this subject over several years and some statistical comparison of the two groups justifies the approach. Undergraduate students performance in other tasks and exam questions, not relevant to the competition, was equivalent to the postgraduate students cohort.

>>>>

<!-- Blinded verssion -->
The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of AB (Computational Statistics and Data Mining, MAST90083, denoted as CSDM), and one at CD University (Statistical Thinking, ETC2420/5242, denoted as ST).

@@

\end{tabular}
\caption{Computational Statistics and Data Mining: Summary statistics of the exam score (out of 100) and the second assignment (out of 10) for the two competition groups.}
\end{center}
\label{tab:Melb_Group_Compare}
\end{table}

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{l l | r | r | r r r r r}\hline
Score & Competition group & N & Mean (SD)   & Min  &  Q1 & Median  &  Q3  & Max \\\hline
Exam   & ST-PG      & 34  & 61.8 (12.2)  &  35.0 & 57.2  &   61.0  & 72.2  & 81.5 \\
       & ST-UG         & 145 & 63.0 (15.4)  &   0.0 & 56.0  &   64.0  & 73.0   & 90.0\\\hline
\end{tabular}
\caption{Statistical Thinking: Summary statistics of the exam score (out of 100) for the two groups.}
\end{center}
\label{tab:Melb_Group_Compare}
\end{table}

@@

```{r Monash_quiz, fig.cap = "\\label{fig:Monash_quiz}Beeswarm plots of undergraduate (ST-UG) and postgraduate (ST-PG) students quizzes during the semester. The scores are very similar, although undergraduates have a longer lower tail.", fig.height=4, fig.width=6}
library(tidyverse)
library(gridExtra)
library(ggbeeswarm)
library(forcats)
quiz <- read_csv(file="data/quiz_scores_by_unit_2420_5242.csv")

@@

#ggplot(quiz, aes(x=unit, y=score, colour=unit)) +
#  geom_beeswarm(alpha=0.5) +
#  facet_wrap(~quiz) + 
#  xlab("") + ylab("Score (%)") +
#  theme(legend.position="none")

@@

ggplot(quiz, aes(x=score, colour=unit, fill=unit)) +
  stat_ecdf(alpha=0.5) +
  facet_wrap(~quiz) + 
  xlab("") + xlab("Score (%)") +
  scale_color_brewer("", palette="Dark2")
  #theme(legend.position="none")

@@

```{r ETC5242, fig.cap = "\\label{fig:ETC5242} Performance for regression question relative to total exam score for students who did and didn't do the regression data competition in Statistical Thinking.", fig.width=4, fig.height=3}
scores_5242 <- read_csv("data/deidentified_scores_5242.csv")
scores_2420 <- read_csv("data/deidentified_scores_2420.csv")

@@

Because the experiment was conducted in the classroom setting as part of the normal teaching of the courses, which imposed limitations on the design. However, performance comparison was enabled in CSDM by a randomized assignment of students to two topic groups, and in ST by using a comparison group. In CSDM, the group sizes were relatively small, approximately 30 students per group. For ST the control group was the undergraduate students that took class. Using undergraduate students as a control group for graduate students may be surprising.  However, the experience of teaching this subject over several years and some statistical comparison of the two groups justifies the approach. Undergraduate students performance in other tasks and exam questions, not relevant to the competition, was equivalent to the postgraduate students cohort.
